\LARGE \bf Planning with Attitude

### 2021-05-05 paper/main.tex
ICRA version
Minor tweaks
Small updates

% Bibliography
\usepackage{biblatex}
\addbibresource{references.bib}

@@

% Make room for more info lines in the \author command 
\author{Brian E Jackson$^{1}$, Kevin Tracy$^{1}$, and Zachary Manchester$^{1}$%
\thanks{Manuscript received: October, 15, 2020; Revised January 2, 2021; Accepted January 4, 2021.}%Use only for final RAL version
\thanks{This paper was recommended for publication by Editor Nancy Amato upon evaluation of the Associate Editor and Reviewers' comments.
This work was supported by a NASA Early Career Faculty Award, NASA JPL, and NSF GFRP Grant No. DGE-1656518.} %Use only for final RAL version
\thanks{$^{1}$All authors are with the Robotics Institute, Carnegie Mellon University, Pittsburgh, PA, USA 
        {\tt\footnotesize brian.jackson@cmu.edu}}%
\thanks{Digital Object Identifier (DOI): see top of this page.}
}

@@

	    \begin{equation} \label{eq:quat_hessian}
            \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) + I_3 \pdv{h}{q}q,
	    \end{equation}
	    where the second term comes from the second derivative of $\varphi(\phi)$.
	    Similar to $G(q)$, this expression is the same (up to a constant scalar factor) for any
        choice of three-parameter attitude representation.

@@

    \begin{equation}
    	\phi_k = ({\nabla r}^T {\nabla r})^{-1} {\nabla r}^T r(q_k).
    \end{equation}
	The update is then applied using the composition for the group:
	\begin{equation}
		\q_{k+1} = \q_k \otimes \varphi(\phi_k).
	\end{equation}
	This ``multiplicative'' Gauss-Newton method is summarized in Algorithm \ref{alg:mgn}.

@@

    We begin by linearizing the dynamics about the reference state and input trajectories, $\bar{x}$ and $\bar{u}$, using
    \eqref{eq:quat_jacobian}. The linearized error dynamics become,
    \begin{equation} \label{eq:linearized_dynamics}
        \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k ,
    \end{equation}
    where \begin{equation}
        \begin{aligned}
            A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \\
            B_k = E(\bar{x}_{k+1})^T \pdv{f}{u}|_{\bar{x}_k,\bar{u}_k},
        \end{aligned}
    \end{equation}
    and $E(x_k) \in \R^{12 \times 13}$ is the error-state Jacobian:
    \begin{equation} \label{eq:state_error_jacobian}
        E(x) = \begin{bmatrix}
            I_3 & & & \\
            & G(q) & & \\
            & & I_3 & \\
            & & & I_3 \\
        \end{bmatrix}\!.
    \end{equation}
    By applying \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} to the nonlinear
    cost functions $\ell$ and \eqref{eq:quat_jacobian} to the nonlinear constraint
    functions $g_k$ and $h_k$, we can calculate the second-order expansion of the cost 
    function:
    \begin{multline}
        \delta \ell(x,u) \approx  \half \delta x^T \ell_{xx} \delta x
            + \half \delta u^T \ell_{uu} \delta u + \delta_u^T \ell_{ux} \delta u \\
            + \ell_x^T \delta x^T + \ell_u^T \delta u.
    \end{multline}
    % \todo{I think you can safely leave out any references to the augmented Lagrangian,
    % etc. and just talk about the cost function.}
    % \begin{equation}
    %     \mathcal{L}_A = \mathcal{L}_N(x_N,\lambda_N,\mu_N) + 
    %         \sum_{k=0}^{N-1} \mathcal{L}_k(x_k,u_k,\lambda_k,\mu_k)
    % \end{equation}
    % where
    % \begin{equation}
    %     \mathcal{L}_k(x,u,\lambda,\mu) = \ell(x,u) + 
    %         (\lambda + \half I_\mu c(x,u))^T c(x,u),
    % \end{equation}
    % with $c(x,u)$ being the concatenation of the constraints $f,g$, and $h$ at a given time step,
    % and $I_\mu$ the penalty matrix.

@@

    \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
        In addition to the straight-forward modifications to the ALTRO algorithm itself, some care must be taken in designing cost functions that are well-suited to unit quaternions. We
        frequently minimize costs that penalize distance from a goal state, e.g. $\half
        (x-x_g)^T Q (x-x_g)$; however, na\"ive substraction of unit quaternions does not respect their group structure, and often results in undesired behavior. Instead, we have found the following cost function, which penalizes the geodesic
        distance between two unit quaternions \cite{Kuffner2004}, to work well in practice:
        \begin{equation} \label{eq:quat_geodesic}
            J_\text{geo} = (1-\abs{q_g^T q}) .
        \end{equation}
        Its gradient and Hessian are,
        \begin{align}
            \nabla J_\text{geo} &= \sign(q_d^T q) q_g^T G(q) \\
            \nabla^2 J_\text{geo} &= \sign(q_d^T q) I_3 q_g^T q ,
        \end{align}
        where $\sign$ denotes the signum function. \added{This cost function is particularly
        useful for rotations since it eliminates the ambiguity arising from the quaternion 
        double-cover of $SO(3)$.}

>>>>

% Bibliography
\usepackage[style=ieee]{biblatex}
\addbibresource{references.bib}

@@

% Make room for more info lines in the \author command 
\author{Brian E Jackson$^{1}$, Kevin Tracy$^{1}$, and Zachary Manchester$^{1}$%
\thanks{Manuscript received: October, 15, 2020; Revised January 2, 2021; Accepted January 4, 2021.}%Use only for final RAL version
\thanks{This paper was recommended for publication by Editor Nancy Amato upon evaluation of the Associate Editor and Reviewers' comments.
This work was supported by a NASA Early Career Faculty Award, NASA JPL, and NSF GFRP Grant No. DGE-1656518.} %Use only for final RAL version
\thanks{$^{1}$All authors are with the Robotics Institute, Carnegie Mellon University, Pittsburgh, PA, USA 
        {\tt\footnotesize brianjackson@cmu.edu}}%
\thanks{Digital Object Identifier (DOI): see top of this page.}
}

@@

	    \begin{equation} \label{eq:quat_hessian}
            \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) - I_3 \pdv{h}{q}q,
	    \end{equation}
	    where the second term comes from the second derivative of $\varphi(\phi)$.
	    Similar to $G(q)$, this expression is the same (up to a constant scalar factor) for any
        choice of three-parameter attitude representation.

@@

    \begin{equation}
    	\phi_k = -({\nabla r}^T {\nabla r})^{-1} {\nabla r}^T r(q_k).
    \end{equation}
	The update is then applied using the composition for the group:
	\begin{equation}
		\q_{k+1} = \q_k \otimes \varphi(\phi_k).
	\end{equation}
	This ``multiplicative'' Gauss-Newton method is summarized in Algorithm \ref{alg:mgn}.

@@

    We begin by linearizing the dynamics about the reference state and input trajectories, $\bar{x}$ and $\bar{u}$, using
    \eqref{eq:quat_jacobian}. The linearized error dynamics become,
    \begin{equation} \label{eq:linearized_dynamics}
        \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k ,
    \end{equation}
    where \begin{equation}
        \begin{aligned}
            A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \\
            B_k = E(\bar{x}_{k+1})^T \pdv{f}{u}|_{\bar{x}_k,\bar{u}_k},
        \end{aligned}
    \end{equation}
    and $E(x_k) \in \R^{13 \times 12}$ is the error-state Jacobian:
    \begin{equation} \label{eq:state_error_jacobian}
        E(x) = \begin{bmatrix}
            I_3 & & & \\
            & G(q) & & \\
            & & I_3 & \\
            & & & I_3 \\
        \end{bmatrix}\!.
    \end{equation}
    By applying \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} to the nonlinear
    cost functions $\ell$ and \eqref{eq:quat_jacobian} to the nonlinear constraint
    functions $g_k$ and $h_k$, we can calculate the second-order expansion of the cost 
    function:
    \begin{multline}
        \delta \ell(x,u) \approx  \half \delta x^T \ell_{xx} \delta x
            + \half \delta u^T \ell_{uu} \delta u + \delta_u^T \ell_{ux} \delta x \\
            + \ell_x^T \delta x^T + \ell_u^T \delta u.
    \end{multline}
    % \todo{I think you can safely leave out any references to the augmented Lagrangian,
    % etc. and just talk about the cost function.}
    % \begin{equation}
    %     \mathcal{L}_A = \mathcal{L}_N(x_N,\lambda_N,\mu_N) + 
    %         \sum_{k=0}^{N-1} \mathcal{L}_k(x_k,u_k,\lambda_k,\mu_k)
    % \end{equation}
    % where
    % \begin{equation}
    %     \mathcal{L}_k(x,u,\lambda,\mu) = \ell(x,u) + 
    %         (\lambda + \half I_\mu c(x,u))^T c(x,u),
    % \end{equation}
    % with $c(x,u)$ being the concatenation of the constraints $f,g$, and $h$ at a given time step,
    % and $I_\mu$ the penalty matrix.

@@

    \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
        In addition to the straight-forward modifications to the ALTRO algorithm itself, some care must be taken in designing cost functions that are well-suited to unit quaternions. We
        frequently minimize costs that penalize distance from a goal state, e.g. $\half
        (x-x_g)^T Q (x-x_g)$; however, na\"ive substraction of unit quaternions does not respect their group structure, and often results in undesired behavior. Instead, we have found the following cost function, which penalizes the geodesic
        distance between two unit quaternions \cite{Kuffner2004}, to work well in practice:
        \begin{equation} \label{eq:quat_geodesic}
            J_\text{geo} = (1-\abs{q_g^T q}) .
        \end{equation}
        Its gradient and Hessian are,
        \begin{align}
            \nabla J_\text{geo} &= -\sign(q_g^T q) q_g^T G(q) \\
            \nabla^2 J_\text{geo} &= \sign(q_g^T q) I_3 q_g^T q ,
        \end{align}
        where $\sign$ denotes the signum function. \added{This cost function is particularly
        useful for rotations since it eliminates the ambiguity arising from the quaternion 
        double-cover of $SO(3)$.}
